Abstractions for Multi-Sorted Substitutions

### 2023-09-08 paper/paper.tex
Update.

  As we intend to have exactly two kit instances, we choose names of
  the form \ACode{x/y}, where \ACode{x} is the name we choose for the
  variable instance, and \ACode{y} is the name we choose for the term
  instance.
  For example the parameter type \ACode{\_∋/⊢\_} will be instanciated
  to \ACode{\ACon{\_∋\_}} for the variable kit, and to \ACode{\AField{\_⊢\_}} for the
  term kit.

@@

  The term kit requires both the variable kit and the
  \ACode{\ACon{Traversal}} to be defined, because shifting a term \ACode{t} with \ACode{\AField{wk}}
  means applying the shifting renaming to the term.
  Hence, we define the term kit in the record module of \ACode{\ACon{Traversal}}:
  \LibCode\KKitTerm

@@

  We start with the operation of applying a map to a term:
  \AppCode\FTraversalOp
  The interesting cases are those with variables and binders:
  \begin{itemize}
  \item 
    In the variable case \ACode{(\ACon{`} x) \ADef{⋯} ϕ}, we first apply the map
    \ACode{ϕ} to the variable \ACode{x}. If \ACode{ϕ} is a renaming,
    we get back a variable and need to apply the variable constructor
    \ACode{\ACon{`\_}} to it. If \ACode{ϕ} is a substitution,
    we get back a term that we can use directly.
    Hence, the use of \ACode{`/id}.
  \item 
    In cases where the operation needs to go under a binder, like
    \ACode{(\ACon{λx} e) \ADef{⋯} ϕ}, we lift the map using \AField{\_↑\_} before
    we apply it to the subterm, to account for the bound variable.
  \end{itemize}

@@

  \subsection{An Examination of Composition}
  With the infrastructure we build in the previous sections, we can
  define the four compositions as followed:
  \ExampleCode\KExFourComps
  The definitions reveal an interesting asymetry, in the sense that
  the first argument of composition guides the implementation: if we
  compose a renaming with any map, then we first apply renaming to the
  the variable, which yields another variable, so we apply the second
  map to that \emph{variable}.
  However, if we compose a substitution with any map, then we first
  apply the substitution to that variable, which yields a term, so we
  need to apply the second map to that \emph{term}.

  With the machinery, we built so far, we can unify the first two and
  last two compositions as followed:
  \ExampleCode\KExTwoComps

  But to unify those two compositions, we need a new abstraction.

  \ExampleCode\KComposition

>>>>

  As we intend to have exactly two \ACode{\ACon{Kit}} instances, we choose names of
  the form \ACode{x/y}, where \ACode{x} is the name we choose for the
  variable instance, and \ACode{y} is the name we choose for the term
  instance.
  For example the parameter type \ACode{\_∋/⊢\_} will be instanciated
  to \ACode{\ACon{\_∋\_}} for the variable kit, and to \ACode{\AField{\_⊢\_}} for the
  term kit.

@@

  The term kit requires both the variable kit and the
  \ACode{\ACon{Traversal}} to be defined, because shifting a term with \ACode{\AField{wk}}
  means applying the shifting renaming to the term.
  Hence, we define the term kit in the record module of \ACode{\ACon{Traversal}}:
  \LibCode\KKitTerm

@@

  We start with the operation of applying a map to a term:
  \AppCode\FTraversalOp
  The interesting cases are those with variables and binders:
  \begin{itemize}
  \item 
    In the variable case \ACode{(\ACon{`} x) \ADef{⋯} ϕ}, we first apply the map
    \ACode{ϕ} to the variable \ACode{x}. If \ACode{ϕ} is a renaming,
    we get back a variable and need to apply the variable constructor
    \ACode{\ACon{`\_}} to it. If \ACode{ϕ} is a substitution,
    we get back a term that we can use directly.
    Hence, the use of \ACode{`/id}.
  \item 
    In cases where the operation needs to go under a binder, like
    \ACode{(\ACon{λx} e) \ADef{⋯} ϕ}, we lift the map using \AField{\_↑\_} before
    we apply it to the subterm to account for the bound variable.
  \end{itemize}

@@

  \subsection{An Examination of Composition}
  To motivate our abstraction for composition, we first look
  at the four compositions individually:
  \ExampleCode\KExFourComps
  The definitions reveal two interesting properties:
  \begin{enumerate}
  \item
    If we compose two maps \ACode{ϕ₁} and \ACode{ϕ₂}, then the
    resulting map is a renaming, iff both \ACode{ϕ₁} and \ACode{ϕ₂} are
    renamings.
    In other words, 
    if \ACode{ϕ₁} is a \ACode{K₁}-map and \ACode{ϕ₂} is a \ACode{K₂}-map,
    then the result is a \ACode{(K₁ ⊔ K₂)}-map, where \ACode{⊔} refers to the
    lattice for \ACode{\{ \ADef{Kᵣ} , \ADef{Kₛ} \}}
    generated by \ACode{\ADef{Kᵣ} $\leq$ \ADef{Kₛ}}.
  \item
    All four compositions first apply \ACode{ϕ₁} to \ACode{x},
    and then apply \ACode{ϕ₂} to the result.
    If \ACode{ϕ₁} is a renaming, this result is another variable, but
    if \ACode{ϕ₁} is a substitution, this result is a term.
  \end{enumerate}
  With the \ACode{\ACon{Kit}} abstraction, we can easily abstract over
  \ACode{ϕ₂} being a renaming or a substitution:
  \ExampleCode\KExTwoComps

  But to abstract over \ACode{ϕ₁}, the \ACode{\ACon{Kit}} abstraction
  is not sufficient: while it allows us to abstract over what we are applying, i.e.
  a renaming or a substitution, it does not allow us to
  abstract over what we are applying it to, i.e. a variable or a term.
  For the latter we have two distinct operations \ACode{\ADef{\_\&\_}}
  and \ACode{\ADef{\_⋯\_}}.

  To fill this gap, we introduce a new abstraction that we call a
  \ACode{\ACon{ComposeKit}},
  which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}, allowing us to
  to define a general composition in the spirit of:
  \ExampleCode\KComposition

  \ACode{\ACon{ComposeKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
  \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-assoc}} are analogous to
  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.
